How hunting experience and prey predictability jointly shape individual foraging and diet specialisation of a predator within an online multiplayer videogame

### 2021-07-08 manuscript/Chapter3-manuscript.Rmd
update manuscript
analysis plan and updated introduction

---
title: "How hunting experience, habitat structure, and prey predictability jointly shape individual specialisation in predator foraging mode in an online videogame"
author: "Maxime Fraser Franco^1^, Francesca Santostefano^1^, Clint D. Kelly^1^, Pierre-Olivier Montiglio^1^"
output: 
  word_document:
    reference_docx: Chapter3-styles.docx
bibliography: references.bib
csl: american-naturalist.csl
---
###### Groupe de Recherche en Écologie Comportementale et Animale (GRECA), Département des Sciences Biologiques, Université du Québec à Montréal, Montréal, QC, Canada

###### ^1^Correspondence: Pavillon des Sciences Biologiques (SB) SB-1805, 141 Avenue du Président-Kennedy, Montréal (Québec), Canada, H2X 1Y4

###### ^1^Corresponding author: maxime(dot)fraser(dot)franco(at)hotmail.com

\newpage
# Abstract {.unnumbered}
This is my abstract paragraph

@@

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [sources]. Recently, ecologists have began to hypothesise that there may be a functional relationship linking individual foraging specialisation to specialisation in resource consumption (i.e. niche specialisation) [sources: Toscano et al. 2016, etc, michalko et gibbons, Kerr, Tinker, Bowen]. One of the postulates of this hypothesis is that predators need to develop specific foraging tactics that are fine tuned to specific prey attributes [sources: Phillips, Woo, etc.]. For instance, predators species often use stereotyped hunting techniques required to capture the type of prey they encounter, such as (dauphins, orcas, etc) [sources : Woo, Estes, tinker, etc]. However, it remains largely unclear how extrinsinc and intrinsinct factors interact to shape these individual-level processes. Therefore, we still have limited knowledge on the ecological consequences of individual foraging and diet specialisation. An important question that remains unanswered is how generalist and specialist individuals can coexist within predator populations. If foraging and resource specialisation can help maintain community stability by promoting species coexistence [sources: bolnick, araujo, wolf, etc], it is crucial that we develop studies that integrate the causes and consequences of these two individual components.

A number of long-term studies have observed similar fitness related measures among generalist and specialist individuals [sources: à partir de Woo]. An emerging explanation is that temporal fluctuations in the predictability of resources (abundance, availability, behaviour) may favor one or the other strategy depending on time scales [sources : Woo, Phillips, Ceia, etc, Courbin, Chang]. The resource-predictability hypothesis advances that when resources are predictable, particularly on short time-scales, individual specialists should benefit from higher fitness returns. The rationale behind this prediction is that individuals should have higher prey delivery rates when they repeatedly employ the same foraging technique as it is easier to assess prey predictability over shorter time-scales. In contrast, individual generalists should be advantaged over longer time scales as resource parameters are expected to fluctuate.

Hunting experience and age have been proposed as major potential drivers of the relationship between foraging specialisation and resource use [sources]. Learning how to hunt is crucial for young predators to reach adulthood and survive.

(learning optimises return rates or foraging success)
(par contre, décomposer l'âge de l'expérience est difficile, et particulièrement chez des espèces sur le terrain puisque ça peut impliquer des techniques invasives)
(parler des études qui ont montré que certains spécialistes étaient meilleurs, mais que c'était confondu par l'âge)

A second important driver of individual differences in predator hunting behaviour is related to habitat-specific features [sources, Kobler, parler de la ideal free distribution?]. Predators are expected to adjust their foraging mode (ambush vs cursorial) to habitat structures that aid/impair (ex. perches, vegetation density) their vision, ability to hide, or energetic/time requirements. For instance, predators who ambush in closed and heterogeneous landscapes often switch to cursorial hunters in open and homogeneous landscapes [sources]. (Donner exemple des orcs)

>>>>

---
title: "How hunting experience and prey predictability jointly shape individual foraging and diet specialisation of a predator within an online multiplayer videogame"
output: 
  word_document:
    reference_docx: Chapter3-styles.docx
bibliography: references.bib
csl: american-naturalist.csl
---

\newpage
# Abstract {.unnumbered}
Key points in this study:

- How experience shapes individual specialisation
  - H1: experience increases the range of behaviors and plasticity by reducing costs to swithing
  - H2: optimizing/refinement process - specialize with time - increase success)

- Prey predictability shapes individual foraging specialization
  - H1: prey predictability is positively correlated to behavioral predictability

- Specialist hunters will fare better on short timescales, when prey predictability is higher
- Generalist hunters will fare better on longer timescales, as they follow prey fluctuations
- Both types of hunters will thus achieve similar hunting success over the sampling period

@@

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [sources]. A potential consequence of among-individual differences in foraging behavior is that they can lead to individual differences in the type of prey captured (i.e. individual niche specialization) [Woo et al., 2008, sources]. There may thus exist a functional relationship linking individual foraging specialisation to specialisation in resource consumption [sources: Toscano et al. 2016, etc, michalko et gibbons, Kerr, Tinker, Bowen]. This would imply that predators need to develop foraging tactics that are fine tuned to specific prey attributes or prey types [sources: Phillips, Woo, etc.]. For instance, predators species often use stereotyped hunting techniques required to capture the type of prey they encounter, such as (dauphins, orcas, etc) [sources : Woo, Estes, tinker, etc]. However, it remains largely unclear how extrinsinc and intrinsinct factors interact to shape these individual differences. Therefore, we still have limited knowledge on how individual foraging specialisation arises, and its ecological and community consequences.

(travailler ceci) An important question that remains unanswered is how can generalist and specialist individuals coexist within predator populations. If foraging and resource specialisation can help maintain community stability by promoting species coexistence [sources: bolnick, araujo, wolf, etc], it is crucial that we develop studies that integrate the causes and consequences of these two individual components.

A number of long-term studies observed that generalist and specialist individuals can achieve similar fitness [sources: à partir de Woo]. An emerging explanation is that temporal fluctuations in the predictability of resources (abundance, availability, behaviour) may favor one or the other strategy depending on time scales [sources : Woo, Phillips, Ceia, etc, Courbin, Chang]. The resource-predictability hypothesis advances that when resources are predictable, particularly on short time-scales, individual specialists should benefit from higher fitness returns. The rationale is that individuals should have higher prey delivery rates when they repeatedly employ the same foraging technique, as it is easier to assess prey predictability over shorter time-scales. In contrast, individual generalists should be advantaged over longer time scales, as resource parameters are expected to fluctuate. [revoir ceia et ramos pour les sources]

Hunting experience has been proposed as a potential major driver of the relationship between foraging specialisation and resource use [sources]. Learning how to hunt is crucial for young predators to reach adulthood and survive.

(learning optimises return rates or foraging success)
(apprentissage, search images, prior information)
(par contre, décomposer l'âge de l'expérience est difficile, et particulièrement chez des espèces sur le terrain puisque ça peut impliquer des techniques invasives)
(parler des études qui ont montré que certains spécialistes étaient meilleurs, mais que c'était confondu par l'âge)
